---
title: "Class6_20230325_DataVisualization_Apichat"
author: "Apichat Photi-A"
date: "2023-03-25"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#attach the libraries
library(socviz)
library(ggplot2)
library(tidyverse)
library(ggrepel)
#install.packages("gapminder")
library(gapminder)
```

```{r}
#build the color palette
model_colors <- RColorBrewer::brewer.pal(3, "Set1")
model_colors
```

```{r}
#Layering geoms - Showing a legend
p0 <- ggplot(data = gapminder, mapping = aes(x = log(gdpPercap), y = lifeExp))
p1 <- p0 + geom_point(alpha = 0.2) + 
  geom_smooth(method = "lm", aes(color = "OLS", fill = "OLS")) + 
  geom_smooth(method = "lm", formula = y ~ splines::bs(x, df = 3), aes(color = "Cubic Spline", fill = "Cubic Spline")) + 
  geom_smooth(method = "loess", aes(color = "LOESS", fill = "LOESS"))
p1

```
```{r}
##make legend pretty
#layering geoms - showing a legend
#assigning a colour from our model_colors palette to each of our model variables
p1 + scale_color_manual(name = "Models", values = model_colors) + scale_fill_manual(name = "Models", values = model_colors) + theme(legend.position = "top")
```

```{r}
gapminder
str(gapminder)
```
```{r}
#create the linear model object 
out <- lm(formula = lifeExp ~ gdpPercap + pop + continent, data = gapminder)
out
str(out)

summary(out)
```

#predict()
```{r}
#make dataframe for predictions
min_gdp <- min(gapminder$gdpPercap)
max_gdp <- max(gapminder$gdpPercap)
med_pop <- median(gapminder$pop)
pred_df <- expand.grid(gdpPercap = (seq(from = min_gdp, to = max_gdp, length.out = 100)), pop = med_pop, 
                       continent = c("Africa", "Americas", "Asia", "Europe", "Oceania")) 

```
```{r}
dim(pred_df)
```
```{r}
##make the prediction
pred_out <- predict(object = out, newdata = pred_df, interval = "predict")
head(pred_out)
```
```{r}
#binding our dataframe
pred_df <- cbind(pred_df, pred_out)
head(pred_df)
str(pred_df)
```

##As an example, we can create a visualization to explore our question “How does per capita GDP affect life expectancy in Europe and Africa?”
```{r}
p <- ggplot(data = subset(pred_df, continent %in% c("Europe", "Africa")), aes(x = gdpPercap, y = fit, ymin = lwr, ymax = upr, color = continent, fill = continent, group = continent))
p + geom_point(data = subset(gapminder, continent %in% c("Europe", "Africa")), 
               aes(x = gdpPercap, 
                   y = lifeExp, color = continent), 
               alpha = 0.5, inherit.aes = FALSE) + 
  geom_line() + 
  geom_ribbon(alpha = 0.2, color = FALSE) + #alpha will show the transparent
  scale_x_log10(labels = scales::dollar)

```
```{r}
p <- ggplot(data = subset(pred_df, continent %in% c("Europe", "Africa")), aes(x = gdpPercap, y = fit, ymin = lwr, ymax = upr, color = continent, fill = continent, group = continent))
p + geom_point(data = subset(gapminder, continent %in% c("Europe", "Africa")), 
               aes(x = gdpPercap, 
                   y = lifeExp), #remove color = continent
               alpha = 0.5, inherit.aes = FALSE) + 
  geom_line() + 
  geom_ribbon(alpha = 0.2, color = FALSE) + #alpha will show the transparent
  scale_x_log10(labels = scales::dollar)

```

How does per capita GDP affect fit(life expectancy) in Asia?
```{r}
p_asia <- ggplot(data = subset(pred_df, continent %in% ("Asia")), aes(x = gdpPercap, y = fit, ymin = lwr, ymax = upr, color = continent, fill = continent, group = continent))
p_asia + geom_point(data = subset(gapminder, continent %in% ("Asia")), 
               aes(x = gdpPercap, 
                   y = lifeExp, color = continent), 
               alpha = 0.5, inherit.aes = FALSE) + 
  geom_line() + 
  geom_ribbon(alpha = 0.2, color = FALSE) + #alpha will show the transparent
  scale_x_log10(labels = scales::dollar)



```

```{r}
#install.packages("broom")
library(broom)
```

```{r}
out_conf <- tidy(out)
out_conf |> round_df()
```

Component level statistics with tidy()
Now we can create a data visualization from this dataframe, the way
that we have been doing all along

```{r}
p <- ggplot(out_conf, mapping = aes(x = term, y = estimate))
p + geom_point() + coord_flip()

```

We can use the confint() function to calculate confidence intervals for our estimates

The tidy() function takes a model object and returns a dataframe
of component level information
- For example, with an added step to round numeric columns to two
decimal places:
```{r}
out_conf <- tidy(out)
out_conf |> round_df() # make it round number
```

plot the results
```{r}
p <- ggplot(out_conf, mapping = aes(x = term, y = estimate))
p + geom_point() + coord_flip()
```
```{r}
out_conf <- tidy(out, conf.int = TRUE)
out_conf %>% round_df()
```

Refine our data
```{r}
out_conf <- subset(out_conf, term %nin% "(Intercept)")
out_conf$nicelabs <- prefix_strip(out_conf$term, "continent")
```

Component level statistics with tidy()
```{r}
#plot our new data
p <- ggplot(out_conf, mapping = aes(x = reorder(nicelabs, estimate), 
                                    y = estimate, 
                                    ymin = conf.low, 
                                    ymax = conf.high))
p + geom_pointrange() + 
  coord_flip() + 
  labs(x = "", 
       y = "OLS Estimate")
```

Observation level statistics with augment()
```{r}
out_aug <- augment(out)
head(out_aug) |> round_df()
```
Plot augmented data 
```{r}
p <- ggplot(data = out_aug, mapping = aes(x = .fitted, y =.resid))
p + geom_point()
```

Model level statistics with glance()
```{r}
glance(out) |> round_df()
```

Using Broom for grouped analysis
```{r}
eu77 <- gapminder |> filter(continent == "Europe", year == 1977)
fit <- lm(lifeExp ~ log(gdpPercap), data = eu77)
summary(fit)
```
Nesting data - let's create a list column
```{r}
out_le <- gapminder |> 
  group_by(continent, year) |>
  nest()
out_le
```
Unnest 
```{r}
out_le |> filter(continent == "Europe" & year == 1977) |>
  unnest()
```

We do this by creating a function that fits our model to a dataframe, then mapping that function to each of our list column’s rows , one at a time
```{r}
# fit_ols <- function(df) {lm(lifeExp ~ log(gdpPercap), data = df)}
# out_le <- gapminder |> 
#   group_by(continent, year) |>
#   nest() |>
#   mutate(model = map(data, fit_ols)) %>%

```
tidy our models
```{r}
# library(dplyr)
# library(purrr)
# fit_ols <- function(df) {lm(lifeExp ~ log(gdpPercap), data = df)}
# 
# out_tidy <- gapminder |> 
#   group_by(continent, year) |> 
#   nest() |>
#   mutate(model = map(data, fit_ols), tidied = map(model, tidy)) |>
#   unnest(tidied) |> 
#   filter(term %nin% "(Intercept)" & continent %nin% "Oceania")

```
Plot our model results
```{r}
# p <- ggplot(data = out_le, 
#             mapping = aes(x = year, 
#                           y = estimate, 
#                           ymin = estimate - 2*std.error, 
#                           ymax = estimate + 2*std.error, 
#                           group = continent, 
#                           color = continent))
# p + geom_pointrange(position = position_dodge(width = 1)) +
#   scale_x_continuous(breaks = unique(gapminder$year)) +
#   theme(legend.position = "top") +
#   labs(x = "Year", y = "Estimate", color = "Continent")
```

